\chapter{INTRODUCTION}

\section{Overview of the Project}

With the rapid advancement of information technology and the widespread availability of the internet, traditional ticket booking systems have gradually evolved into digital platforms. Transportation services, especially bus travel, continue to play a vital role in connecting cities and towns, particularly in regions where buses remain the most economical and accessible mode of transport. However, conventional bus ticket booking methods often involve manual processes, long waiting times, lack of real-time seat availability, and limited flexibility for passengers.

The Bus Ticket Booking System is developed to address these challenges by providing a user-friendly, efficient, and web-based platform for booking bus tickets. The system enables users to search for buses based on source, destination, and travel date, view detailed information about routes, timings, bus types, and fares, and select seats according to their preferences. This digital approach significantly reduces manual effort and enhances user convenience.

The project primarily focuses on the frontend development of the application using React.js and modern web technologies. The application is implemented as a single-page application (SPA) to ensure smooth navigation and faster performance. It includes key features such as user registration and login, bus search, seat selection, passenger detail entry, checkout, and invoice generation.

Additionally, an admin module is provided to manage bus details, schedules, routes, and fare structures. Although backend and real-time payment integration are not included, the project effectively demonstrates a scalable foundation for a complete online bus booking system.

\section{Objectives of the Project}

The primary objectives of this project are as follows:

\begin{itemize}
    \item To design a user-friendly and responsive frontend for an online bus booking system.
    \item To allow users to search buses based on source, destination, and travel date.
    \item To implement seat selection and passenger detail entry.
    \item To provide secure login and registration functionality.
    \item To restrict checkout access to logged-in users.
    \item To provide an admin interface for managing buses, routes, schedules, and fares.
    \item To demonstrate frontend development concepts using React.js.
\end{itemize}

\section{Scope of the Project}

The scope of this project is limited to frontend development. The system includes:

\begin{itemize}
    \item User registration and login
    \item Bus search and filtering
    \item Seat selection and passenger information management
    \item Checkout and payment simulation
    \item Admin dashboard for bus management
\end{itemize}

Backend features such as database connectivity, real payment gateways, and server-side authentication are not implemented and are considered future enhancements.
